---
title: "microclimate_troubleshoot"
output: html_document
---

## Setup

```{r}
librarian::shelf(tidyverse, lubridate, fuzzyjoin, zoo, janitor)

# load data files
## I refer to the re-exported data from the .hobo files as "raw", and the processed data as "combined" (it comes from the "combined" hob data files)

# reference combined file (this one should be good)
combined_feb_in = read_csv("hobo_inquiry_data/PA_combined_Feb24-Jun24.csv")
# discrepant combined file (this one seems to show illuminance discrepancies)
combined_jun_in = read_csv("hobo_inquiry_data/PA_combined_Jun24-Oct24.csv")

# re-exported from hoboware: feb-jun 24 & jun - oct 24 data file (one combined file)
raw_in = read_csv("hobo_inquiry_data/Phelps_shade_25cm_feb24_oct24_from_hobo.csv", skip = 1)
# split into feb & jun for simplicity
raw_feb_in = raw_in %>%
  filter(`#` <= 6006)

raw_jun_in = raw_in %>%
  filter(`#` >= 6010)

```

# clean files
```{r}
# clean combined files
combined_feb_clean = combined_feb_in %>%
  clean_names() %>%
  filter(site == "Phelps",
         location == "shade",
         height == "25cm") %>%
  mutate(date_time_combined = parse_date_time(`date_time`, c("ymd HMS"))) %>%
  rename(temperature_combined = temperature,
         intensity_lux_combined = intensity_lux)

combined_jun_clean = combined_jun_in %>%
  clean_names() %>%
  filter(site == "Phelps",
         location == "shade",
         height == "25cm") %>%
  mutate(date_time_combined = parse_date_time(`date_time`, c("ymd HMS"))) %>%
  rename(temperature_combined = temperature,
         intensity_lux_combined = intensity_lux)
  
# clean raw files
raw_feb_clean = raw_feb_in %>%
  clean_names() %>%
  mutate(date_time_raw = parse_date_time(`date_time_gmt_05_00`, c("mdy IMSp"))) %>%
  rename(temperature_raw = temp_c_lgr_s_n_21359531_sen_s_n_21359531_lbl_pa_phelps_temp_air_middle,
         intensity_lux_raw = intensity_lux_lgr_s_n_21359531_sen_s_n_21359531_lbl_pa_phelps_light_air_middle)

raw_jun_clean = raw_jun_in %>%
  clean_names() %>%
  mutate(date_time_raw = parse_date_time(`date_time_gmt_05_00`, c("mdy IMSp"))) %>%
  rename(temperature_raw = temp_c_lgr_s_n_21359531_sen_s_n_21359531_lbl_pa_phelps_temp_air_middle,
         intensity_lux_raw = intensity_lux_lgr_s_n_21359531_sen_s_n_21359531_lbl_pa_phelps_light_air_middle)

# the minutes/seconds of the raw_jun (hh:17:09 & hh:47:09) and combined_jun (hh:13:00 & hh:43:00) data don't line up. A clue? Is this even the same data?

# fuzzy join in time
feb_joined = raw_feb_clean %>%
  select(date_time_raw,
         temperature_raw,
         intensity_lux_raw) %>%
  fuzzy_full_join(combined_feb_clean,
                  by = c("date_time_raw" = "date_time_combined"),
                  match_fun = \(x, y) abs(as.numeric(x - y)) < 15)  # join to nearest in time, within a 15-minute window

jun_joined = raw_jun_clean %>%
  select(date_time_raw,
         temperature_raw,
         intensity_lux_raw) %>%
  fuzzy_full_join(combined_jun_clean,
                  by = c("date_time_raw" = "date_time_combined"),
                  match_fun = \(x, y) abs(as.numeric(x - y)) < 15)  # join to nearest in time, within a 15-minute window
```

observation: the times are off even down to the minute/second... is this even the same data?

# daily illuminance diagnostic
```{r}
all_bind = bind_rows(feb_joined %>%
                       mutate(src = "Reference Feb24 - Jun24"),
                     jun_joined %>%
                       mutate(src = "Discrepant Jun24 - Oct24")) %>%
  mutate(time_combined = as_hms(date_time_combined),
         time_raw = as_hms(date_time_raw))

ggplot(all_bind, aes(x = time_combined, y = log1p(intensity_lux_combined), color = src)) +
  geom_point(alpha = 0.1) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  scale_x_time(breaks = as_hms(c("04:00:00", "08:00:00", "12:00:00", "16:00:00", "20:00:00")),
                   labels = c("04:00", "08:00", "12:00", "16:00", "20:00")) +
  labs(title = "Combined-file illuminance data")

ggplot(all_bind, aes(x = time_raw, y = log1p(intensity_lux_raw), color = src)) +
  geom_point(alpha = 0.1) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  scale_x_time(breaks = as_hms(c("04:00:00", "08:00:00", "12:00:00", "16:00:00", "20:00:00")),
                   labels = c("04:00", "08:00", "12:00", "16:00", "20:00")) +
  labs(title = "Raw illuminance data")
```

We see a discrepancy in the timing of dusk and dawn between the reference timeseries and the discrepant one, using the data from the combined hobo files. Running the same comparison using the re-exported "raw" data, we see that they agree much better. So something in how the discrepant time series was processed is resulting in this error.

# simple visual checks

How correlated are raw and combined time series, for reference and discrepant data?
```{r}
# compare illuminance
ggplot(feb_joined, aes(x = log1p(intensity_lux_raw), y = log1p(intensity_lux_combined))) +
  geom_point(alpha = 0.2) +
  labs(title = "Reference intensity data (Feb24 - Jun24)")

ggplot(jun_joined, aes(x = log1p(intensity_lux_raw), y = log1p(intensity_lux_combined))) +
  geom_point(alpha = 0.2) +
  labs(title = "Discrepant intensity data (Jun24 - Oct24)")


# compare temperatures
ggplot(feb_joined, aes(x = temperature_raw, y = temperature_combined)) +
  geom_point(alpha = 0.2) +
  labs(title = "Reference temperature data (Feb24 - Jun24)")

ggplot(jun_joined, aes(x = temperature_raw, y = temperature_combined)) +
  geom_point(alpha = 0.2) +
  labs(title = "Discrepant temperature data (Jun24 - Oct24)")


```
no apparent correlation for the discrepant (jun) data
perfect alignment for the reference (feb) data

# q-q plots
Are raw and combined time series from the same distributions, for reference and discrepant data?
```{r}
gg_qq_empirical <- function(a, b, n = 100) {
  probs <- seq(0, 1, length.out = n)
  df <- data.frame(x = quantile(a, probs, na.rm = TRUE),
                   y = quantile(b, probs, na.rm = TRUE))
  ggplot(df, aes(x = x, y = y)) +
    geom_point() +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") +
    coord_equal() +
    labs(x = deparse(substitute(a)), y = deparse(substitute(b)))
}

feb_joined_inner = feb_joined %>%
  filter(!is.na(date_time_raw),
         !is.na(date_time_combined))

jun_joined_inner = jun_joined %>%
  filter(!is.na(date_time_raw),
         !is.na(date_time_combined)) %>%
  group_by(date_time_raw) %>%
  slice_head(n = 1) %>%
  ungroup()

gg_qq_empirical(log1p(feb_joined_inner$intensity_lux_raw), log1p(feb_joined_inner$intensity_lux_combined), n = 50) +
  labs(title = "Reference intensity data (Feb24 - Jun24)")
gg_qq_empirical(feb_joined_inner$temperature_raw, feb_joined_inner$temperature_combined, n = 50) +
  labs(title = "Discrepant intensity data (Jun24 - Oct24)")

gg_qq_empirical(log1p(jun_joined_inner$intensity_lux_raw), log1p(jun_joined_inner$intensity_lux_combined), n = 50) +
  labs(title = "Reference temperature data (Feb24 - Jun24)")
gg_qq_empirical(jun_joined_inner$temperature_raw, jun_joined_inner$temperature_combined, n = 50) +
  labs(title = "Discrepant temperature data (Jun24 - Oct24)")

```

Both combined and raw data from reference and discrepant datasets appear to be from the same distributions (for both temperature and illuminance). This suggests a shift in time...

# cross correlation analysis
```{r}
ccor_func = function(datetime_a, var_a, datetime_b, var_b, max_h = 24) {
  overlap_start <- max(min(datetime_a), min(datetime_b))
  overlap_end   <- min(max(datetime_a), max(datetime_b))
  
  # Define a regular time grid at your sampling interval
  sample_interval <- 30 * 60  # in seconds 
  time_grid <- seq(overlap_start, overlap_end, by = sample_interval)
  
  # Interpolate both series onto the common grid
  zoo_a <- zoo(var_a, datetime_a)
  zoo_b <- zoo(var_b, datetime_b)
  
  reg_a <- na.approx(zoo_a, xout = time_grid, na.rm = FALSE)
  reg_b <- na.approx(zoo_b, xout = time_grid, na.rm = FALSE)
  
  # Cross-correlate
  ccf_result <- ccf(coredata(reg_a), coredata(reg_b), 
                    lag.max = 2 * max_h,  # n samples x 30min / sample
                    na.action = na.pass,
                    plot = TRUE)
  
  # Find the lag of maximum correlation
  best_lag_steps <- ccf_result$lag[which.max(ccf_result$acf)]
  best_lag_seconds <- best_lag_steps * sample_interval
  best_lag_minutes <- best_lag_seconds / 60
  
  cat("Best lag:", best_lag_minutes, "minutes\n")
  cat("Max correlation:", max(ccf_result$acf), "\n")
}

ccor_func(feb_joined_inner$date_time_raw, feb_joined_inner$intensity_lux_raw,
          feb_joined_inner$date_time_combined, feb_joined_inner$intensity_lux_combined,
          max_h = 24 * 2)

ccor_func(feb_joined_inner$date_time_raw, feb_joined_inner$temperature_raw,
          feb_joined_inner$date_time_combined, feb_joined_inner$temperature_combined,
          max_h = 24 * 2)

ccor_func(jun_joined_inner$date_time_raw, jun_joined_inner$intensity_lux_raw,
          jun_joined_inner$date_time_combined, jun_joined_inner$intensity_lux_combined,
          max_h = 24 * 2)

ccor_func(jun_joined_inner$date_time_raw, jun_joined_inner$temperature_raw,
          jun_joined_inner$date_time_combined, jun_joined_inner$temperature_combined,
          max_h = 24 * 2)

```
Subtract 960 minutes from date_time_combined, and compare time series
```{r}
jun_shift = jun_joined_inner %>%
  mutate(date_time_combined_shifted = date_time_combined - minutes(955) - seconds(51))

# explicit join (not fuzzy)
jun_compare = jun_shift %>%
  select(date_time_raw,
         intensity_lux_raw,
         temperature_raw) %>%
  full_join(jun_shift %>%
                    select(date_time_combined,
                           date_time_combined_shifted,
                           intensity_lux_combined,
                           temperature_combined),
                  by = c("date_time_raw" = "date_time_combined_shifted"),
            keep = TRUE)

# compare illuminance
ggplot(jun_compare, aes(x = log1p(intensity_lux_raw), y = log1p(intensity_lux_combined))) +
  geom_point(alpha = 0.2) +
  labs(title = "Discrepant data (Jun24 - Oct24)")


# compare temperatures
ggplot(jun_compare, aes(x = temperature_raw, y = temperature_combined)) +
  geom_point(alpha = 0.2) +
  labs(title = "Discrepant data (Jun24 - Oct24)")
```

Alignment!

But where does this errant -955 minute 9 second shift come from? (maybe the 9 second shift is due to rounding?)