---
title: "legacy_panama_qpcr_wrangling"
format: html
editor: source
---

## Load Packages

```{r}

if(!require(librarian)){
  install.packages(librarian)
  library(librarian)
}

shelf(tidyverse, here, DBI, RPostgres, janitor, readxl, parsedate)

```

## Read in all .xlxs files

```{r}

file_list <- gsub("\\.xlsx$","", list.files(here("panama_data")))


for (i in file_list){
  
  assign(i, read_excel(here("data", paste(i, ".xlsx", sep=""))))
  
}


```

Columns: swab_id, field_protocol, qPCR_protocol, qPCR_instrument, standard, ITS1_region_replicates, extraction_chemistry, replicate, negative_control, zoospore_eq/average_zoospores, ITS1_copies

## Atelopus

```{r}

clean_atelopus <- atelopus %>% 
  clean_names() %>% 
  select(!c(sample_type, site, species, home_base, preservative, notes_7, status, 
            location_of_sample)) %>% 
  rename(comments = notes_19,
         bd_swab_id = frog_id,
         average_zoospores = swab_result,
         average_copy_number = quantity) %>% 
  mutate(extraction_kit = str_to_lower(str_replace_all(extraction_kit, " ", "_")),
         extraction_kit = str_replace(extraction_kit, "-", "_"),
         ct = if_else(is.na(ct), 0, ct),
         average_copy_number = if_else(is.na(average_copy_number), 0, average_copy_number),
         average_zoospores = if_else(is.na(average_zoospores), 0, average_zoospores),
         bd_swab_id = str_to_lower(str_replace_all(bd_swab_id, "-", "_")),
         standard = "pisces",
         master_mix = "bioline",
         result = if_else(result == "positive", "p", result),
         result = if_else(result == "negative", "n", result),
         replicate = "singley",
         qpcr_machine = "quantstudio_3")


```

## fall_2012 - extraction kit?

```{r}

clean_fall_2012 <- fall_2012 %>% 
  clean_names() %>% 
  select(!c(site, species)) %>% 
  rename(comments = notes,
         bd_swab_id = swab,
         extraction_plate_id = extraction_plate,
         pcr_plate_id = pcr_plate) %>% 
  mutate(bd_swab_id = str_replace_all(bd_swab_id, "-", "_"),
         average_zoospores = if_else(is.na(average_zoospores), 0, average_zoospores),
         bd_swab_id = str_remove_all(bd_swab_id, "\\*"),
         standard = "pisces",
         master_mix = "taqman",
         replicate = "triplicate",
         qpcr_machine = "applied_biosystems_7500",
         extraction_kit = "qiagen_dneasy")
  

```

## qpcr_2014

```{r}

clean_2014 <- qpcr_2014 %>% 
  clean_names() %>% 
  select(!c(site, species)) %>% 
  rename(bd_swab_id = id,
         average_zoospores = swab_result,
         comments = additional_notes,
         pcr_plate_id = pcr_plate,
         extraction_plate_id = extraction_plate,
         ct = ct_vlues) %>% 
  mutate(result = if_else(result == "0", "n,n,n", result),
         bd_swab_id = str_to_lower(str_replace_all(bd_swab_id, "-", "_")),
         ct = if_else(is.na(ct), 0, ct),
         average_copy_number = if_else(is.na(average_copy_number), 0, average_copy_number),
         average_zoospores = if_else(is.na(average_zoospores), 0, average_zoospores),
         standard = "pisces",
         master_mix = "taqman",
         replicate = "triplicate",
         qpcr_machine = "abi_7500",
         extraction_kit = "qiagen_dneasy")

```

## winter_2013

```{r}

clean_winter_2013 <- winter_2013 %>% 
  clean_names() %>% 
  select(!c(site, species)) %>% 
  rename(bd_swab_id = swab,
         average_zoospores = swab_result,
         comments = additional_notes,
         pcr_plate_id = pcr_plate,
         extraction_plate_id = extraction_plate,
         ct = ct_vlues) %>% 
  mutate(ct = if_else(is.na(ct), 0, ct),
         average_copy_number = if_else(is.na(average_copy_number), 0, average_copy_number),
         average_zoospores = if_else(is.na(average_zoospores), 0, average_zoospores),
          bd_swab_id = str_to_lower(str_replace_all(bd_swab_id, "-", "_")),
         standard = "pisces",
         master_mix = "taqman",
         replicate = "triplicate",
         qpcr_machine = "abi_7500",
         extraction_kit = "qiagen_dneasy")

```

## summer_2013

```{r}

clean_summer_2013 <- summer_2013 %>% 
  clean_names() %>% 
  select(!c(site, species, location)) %>% 
  rename(bd_swab_id = swab,
         average_zoospores = swab_result,
         comments = additional_notes,
         pcr_plate_id = pcr_plate,
         extraction_plate_id = extraction_plate,
         ct = ct_number) %>% 
  mutate(ct = if_else(is.na(ct), 0, ct),
         average_copy_number = if_else(is.na(average_copy_number), 0, average_copy_number),
         average_zoospores = if_else(is.na(average_zoospores), 0, average_zoospores),
          bd_swab_id = str_to_lower(str_replace_all(bd_swab_id, "-", "_")),
         standard = "pisces",
         master_mix = "taqman",
         replicate = "triplicate",
         qpcr_machine = "abi_7500",
         extraction_kit = "qiagen_dneasy")

```

## qpcr_2015

```{r}

clean_2015 <- qpcr_2015 %>% 
  clean_names() %>% 
  select(!c(site, species)) %>% 
  rename(bd_swab_id = frog_id,
         average_zoospores = swab_result,
         comments = additional_notes,
         ct = mean_ct_value,
         average_copy_number = mean_copy_number) %>% 
  mutate(ct = if_else(is.na(ct), 0, ct),
         average_copy_number = if_else(is.na(average_copy_number), 0, average_copy_number),
         average_zoospores = if_else(is.na(average_zoospores), 0, average_zoospores),
         result = if_else(result == "0", "n,n,n", result),
         bd_swab_id = str_to_lower(str_replace_all(bd_swab_id, "-", "_")),
         standard = "pisces",
         master_mix = "taqman",
         replicate = "triplicate",
         qpcr_machine = "quantstudio_3",
         extraction_kit = "qiagen_dneasy")
  

```

## qpcr_2016

```{r}

clean_2016 <- qpcr_2016 %>% 
  clean_names() %>% 
  select(!c(site, species)) %>% 
  rename(bd_swab_id = frog_id,
         average_zoospores = swab_result,
         comments = additional_notes,
         ct = mean_ct_value,
         average_copy_number = mean_copy_number)%>% 
  mutate(ct = if_else(is.na(ct), 0, ct),
         average_copy_number = if_else(is.na(average_copy_number), 0, average_copy_number),
         average_zoospores = if_else(is.na(average_zoospores), 0, average_zoospores),
         result = if_else(result == "0", "n,n,n", result),
         bd_swab_id = str_to_lower(str_replace_all(bd_swab_id, "-", "_")),
         standard = "pisces",
         master_mix = "bioline",
         replicate = "triplicate",
         qpcr_machine = "quantstudio_3",
         extraction_kit = "qiagen_dneasy")


```

## js086-88 missing IDs - NEED to calculate out average zoospore

```{r}

js86 <- js086 %>% 
  slice(c(7:199)) %>% 
  row_to_names(row_number = 1) %>% 
  clean_names()

js87 <- js087 %>% 
  slice(c(7:103)) %>% 
  row_to_names(row_number = 1) %>% 
  clean_names()

js88 <- js088 %>% 
  slice(c(7:103)) %>% 
  row_to_names(row_number = 1) %>% 
  clean_names()

js_samps <- plyr::rbind.fill(js86, js87, js88) %>% 
  select(sample_name, ct_mean, quantity_mean, comments) %>% 
  filter(!str_detect(sample_name, "Pisces"),
         !str_detect(sample_name, "NEG"),
         !str_detect(sample_name, "POS")) %>% 
  mutate(sample_name = str_replace_all(str_to_lower(sample_name), "-", "_"),
         ct_mean = as.numeric(ct_mean),
         quantity_mean = as.numeric(quantity_mean),
         ct_mean = round(if_else(is.na(ct_mean), 0, ct_mean), 2),
         quantity_mean = round(if_else(is.na(quantity_mean), 0, quantity_mean), 2)) %>% 
  rename(bd_swab_id = sample_name,
         ct = ct_mean,
         quantity = quantity_mean) %>% 
  mutate(result = if_else(!ct == 0, "p", "n")) %>% 
  group_by(bd_swab_id) %>% 
  reframe(ct = mean(ct),
          average_copy_number = mean(quantity),
          result = aggregate(result ~ bd_swab_id, FUN = paste, collapse = ",")) %>% 
  unnest() %>% 
  select(!c(bd_swab_id1)) %>% 
  mutate(standard = "pisces",
         master_mix = "taqman",
         replicate = "quadruplicates",
         qpcr_machine = "abi_7500",
         extraction_kit = "qiagen_dneasy",
         pcr_plate_id = "js086-088",
         average_zoospores = (average_copy_number * 400),
         comments = "CAUTION average_zoospores values calculated estimate of a 400 dillution factor - CZ")

```

## 2019_qpcr

```{r}

clean_2019 <- qpcr_2019 %>% 
  clean_names() %>% 
  select(!c(well, well_position, site, region_code, site_code, species, ipc, bd_pos_neg, extraction_tech, extraction_lab, q_pcr_tech, q_pcr_lab, ext_pos_quality, ext_neg_quality, sample_type, rerun)) %>% 
  rename(bd_swab_id = sample_name,
         ct = bd_ct_mean,
         average_copy_number = bd_quantity_mean,
         extraction_plate_id = extraction_plate,
         pcr_plate_id = q_pcr_plate) %>% 
  mutate(standard = "pisces",
         master_mix = "bioline",
         replicate = "triplicate",
         qpcr_machine = "quantstudio_3",
         extraction_kit = "qiagen_dneasy",
         ct = if_else(is.na(ct), 0, ct),
         average_copy_number = if_else(is.na(average_copy_number), 0, average_copy_number),
         average_zoospores = (40*average_copy_number))
  

```

## read in qpcr_lab and clean

```{r}

labs <- read_csv(here("data_csv", "2012_2017_bdswab_search_clz.csv"))

clean_labs <- labs %>% 
  select(bd_swab_id, `Who did the qPCR?`) %>% 
  rename(qpcr_lab = `Who did the qPCR?`) %>% 
  mutate(qpcr_lab = if_else(qpcr_lab == "Cori's lab - Pitt", "rz_lab", qpcr_lab),
         qpcr_lab = if_else(qpcr_lab == "Jamie's lab", "jv_lab", qpcr_lab),
         qpcr_lab = if_else(qpcr_lab == "??", NA, qpcr_lab),
         qpcr_lab = if_else(qpcr_lab == "Matt Robak - Tulane", "rz_lab", qpcr_lab),
         qpcr_lab = if_else(qpcr_lab == "Julia Sonn - Tulane", "rz_lab", qpcr_lab),
         extract_lab = qpcr_lab)

```

## Bind all tables together

```{r}

panama_qpcr <- plyr::rbind.fill(clean_2014, clean_2015, clean_2016, clean_atelopus, clean_fall_2012, clean_winter_2013, clean_summer_2013, 
                                js_samps, clean_2019) %>% 
  rename(average_ct = ct,
         qpcr_plate_id = pcr_plate_id,
         qpcr_date = q_pcr_date) %>% 
  mutate(average_zoospores = round(average_zoospores, 2),
         average_copy_number = round(average_copy_number, 2),
         average_ct = round(average_ct, 2),
         across(c(c(bd_swab_id:qpcr_plate_id), c(comments:extraction_kit)), str_to_lower),
         average_copy_number = if_else(average_copy_number == 0, NA, average_copy_number),
         average_ct = if_else(average_ct == 0, NA, average_ct),
         average_zoospores = if_else(average_zoospores == 0, NA, average_zoospores)) %>% 
  left_join(clean_labs, by = c("bd_swab_id")) %>% 
  mutate(qpcr_lab = if_else(str_detect(bd_swab_id, "^18"), "rz_lab", qpcr_lab),
         qpcr_lab = if_else(str_detect(bd_swab_id, "^19"), "rz_lab", qpcr_lab),
         extract_lab = if_else(str_detect(bd_swab_id, "^19"), "rz_lab", extract_lab),
         extract_lab = if_else(str_detect(bd_swab_id, "^18"), "rz_lab", extract_lab),
         qpcr_lab = if_else(is.na(qpcr_lab), "rz_lab", qpcr_lab),
         extract_lab = if_else(is.na(extract_lab), "rz_lab", extract_lab),
         swab_type = "mw113",
         total_volume_uL = if_else(master_mix == "taqman", 50, 200),
         dillution_factor = if_else(row_number() > 2942, "1:1", "1:10"), # very dangerous line of code. Make sure nothin above changes
         volume_template_dna_uL = 5)

# missing_labs <- panama_qpcr %>% 
#   filter(is.na(extract_lab)) %>% 
#   write_csv(here("missing_labs.csv"))

write_csv(panama_qpcr, here("clean_qpcr", "panama_qpcr.csv"))

```

## find missing IDs still

```{r}

# all_ids <- read_csv(here("data_csv", "2012_2017_bdswab_search_clz.csv"))
# 
# missing_ids <-  all_ids %>%
#   anti_join(panama_qpcr, by = c("bd_swab_id"))
# 
# missing_ids <- missing_ids[-(90:187),]
# 
# write_csv(missing_ids, here("data_csv", "remaining_missing_ids.csv"))
# 
# 
# 
# plates <- panama_qpcr %>%
#   select(bd_swab_id, extraction_plate_id, pcr_plate_id) %>%
#   write_csv(here("data_csv", "extraction_ids.csv"))
  

```
