---
title: "frog_temp_microclimate_query"
output: html_document
---

# pull data for microclimate & frog body temp where we have both

Considerations
 - Loggers record at regular intevals generally not aligned with capture times. Fuzzy matching needed.
 - Not all captures have an associated time_of_capture, next best sit start_time and end_time of survey (usually ~1h range)
 - How do we want to 

```{r}

# librarian downloads, if not already downloaded, and reads in needed packages
librarian::shelf(tidyverse, dbplyr, here, RPostgres, DBI, RIBBiTR-BII/ribbitrrr, data.table)

# connect to database
dbcon = hopToDB("ribbitr")

```

# table pointers
```{r}
# survey data
db_bd = tbl(dbcon, Id("survey_data", "bd_qpcr_results"))
db_sample = tbl(dbcon, Id("survey_data", "sample"))
db_capture = tbl(dbcon, Id("survey_data", "capture"))
db_env = tbl(dbcon, Id("survey_data", "environmental"))
db_survey = tbl(dbcon, Id("survey_data", "survey"))
db_visit = tbl(dbcon, Id("survey_data", "visit"))
db_site = tbl(dbcon, Id("survey_data", "site"))
db_region = tbl(dbcon, Id("survey_data", "region"))
db_country = tbl(dbcon, Id("survey_data", "country"))

# microclimate data
db_ts_temp = tbl(dbcon, Id("microclimate_data", "ts_temperature"))
db_sensor = tbl(dbcon, Id("microclimate_data", "sensor"))
db_logger = tbl(dbcon, Id("microclimate_data", "logger"))
```

# joining data
```{r}
# microclimate data
logger_site_date = db_ts_temp %>%
  left_join(db_sensor, by = "sensor_id") %>%
  left_join(db_logger, by = "logger_id") %>%
  group_by(site_id) %>%
  summarise(ts_start_date = date(min(timestamp_utc, na.rm = TRUE)),
            .groups = "drop")

# survey data, filtered to captures at sites with hobo loggers, placed before capture
data_capture = db_capture %>%
  left_join(db_survey, by = "survey_id") %>%
  left_join(db_visit, by = "visit_id") %>%
  left_join(db_site, by = "site_id") %>%
  left_join(db_region, by = "region_id") %>%
  left_join(db_country, by = "country_id") %>%
  inner_join(logger_site_date, by = "site_id") %>%
  filter(date >= ts_start_date,
         !is.na(body_temp_c))

# list of unique visit ids
visit_ids = data_capture %>%
  select(visit_id) %>%
  distinct() %>%
  pull(visit_id)

# environmental surveys, filtered to visit ids
data_env = db_env %>%
  left_join(db_survey, by = "survey_id") %>%
  left_join(db_visit, by = "visit_id") %>%
  filter(visit_id %in% visit_ids,
         !is.na(air_temp_c) | !is.na(water_temp_c)) %>%
  group_by(visit_id) %>%
  slice_min(order_by = "air_time", n = 1) %>%
  ungroup() %>%
  select(visit_id,
         air_temp_c,
         air_time,
         water_temp_c,
         water_time)

# bd qpcr results, summarised across replicates
data_bd = db_bd %>%
  inner_join(db_sample, by = "sample_id") %>%
  group_by(sample_id, sample_name_bd, capture_id) %>%
  summarise(bd_qpcr_replicates = n(),
            bd_detected = any(detected, na.rm = TRUE),
            bd_mean_its1_copies_per_swab = mean(bd_its1_copies_per_swab, na.rm = TRUE),
            .groups = "drop")

# bring data queries together, select vars of interest, collect data
data_frog = data_capture %>%
  left_join(data_env, by = "visit_id") %>%
  left_join(data_bd, by = "capture_id") %>%
  rename(survey_start_time = start_time,
         survey_end_time = end_time) %>%
  select(capture_id,
         taxon_capture,
         svl_mm,
         body_mass_g,
         sex,
         sample_name_bd,
         bd_qpcr_replicates,
         bd_detected,
         bd_mean_its1_copies_per_swab,
         body_temp_c,
         substrate_temp_c,
         microhabitat_type,
         microhabitat_wet,
         microhabitat_detailed,
         time_of_capture,
         timestamp_of_capture_utc,
         air_temp_c,
         air_time,
         water_temp_c,
         water_time,
         survey_start_time,
         start_timestamp_utc,
         survey_end_time,
         end_timestamp_utc,
         date,
         site_id,
         site,
         region,
         country,
         time_zone) %>%
  collect()

# export
write_csv(data_frog, here("staging", paste0("frog_temp_", today(), ".csv")))

```

# summarize data
```{r}
frog_summary = data_frog %>%
  mutate(year = year(date)) %>%
  group_by(country, region, year) %>%
  count()

```

# sn inquiry
```{r}
df = data_frog %>%
  filter(region == "california")
  
lsd = logger_site_date %>%
  left_join(db_site, by = "site_id") %>%
  left_join(db_region, by = "region_id") %>%
  left_join(db_country, by = "country_id") %>%
  filter(region == "california") %>%
  collect()

dc = db_capture %>%
  left_join(db_survey, by = "survey_id") %>%
  left_join(db_visit, by = "visit_id") %>%
  left_join(db_site, by = "site_id") %>%
  left_join(db_region, by = "region_id") %>%
  left_join(db_country, by = "country_id") %>%
  filter(region == "california",
         !is.na(body_temp_c)) %>%
  collect()

dcs = dc %>%
  group_by(site) %>%
  count() %>%
  left_join(lsd %>%
              select(site,
                     ts_start_date), by = "site") %>%
  rename(captures_with_temp = n,
         hobo_start_date = ts_start_date)

# write_csv(dcs, here("staging", paste0("sn_sites_loggers_frog_temp_", today(), ".csv")))


```

# sn all body-temp captures
```{r}
# microclimate data
logger_site_date = db_ts_temp %>%
  left_join(db_sensor, by = "sensor_id") %>%
  left_join(db_logger, by = "logger_id") %>%
  group_by(site_id) %>%
  summarise(ts_start_date = date(min(timestamp_utc, na.rm = TRUE)),
            .groups = "drop")

data_capture = db_capture %>%
  left_join(db_survey, by = "survey_id") %>%
  left_join(db_visit, by = "visit_id") %>%
  left_join(db_site, by = "site_id") %>%
  left_join(db_region, by = "region_id") %>%
  left_join(db_country, by = "country_id") %>%
  # left_join(logger_site_date, by = "site_id") %>%
  filter(
    # date >= ts_start_date,
    !is.na(body_temp_c),
    region == "california")

# list of unique visit ids
visit_ids = data_capture %>%
  select(visit_id) %>%
  distinct() %>%
  pull(visit_id)

# environmental surveys, filtered to visit ids
data_env = db_env %>%
  left_join(db_survey, by = "survey_id") %>%
  left_join(db_visit, by = "visit_id") %>%
  filter(visit_id %in% visit_ids,
         !is.na(air_temp_c) | !is.na(water_temp_c)) %>%
  group_by(visit_id) %>%
  slice_min(order_by = air_time, n = 1) %>%
  ungroup() %>%
  select(visit_id,
         air_temp_c,
         air_time,
         water_temp_c,
         water_time)

# bd qpcr results, summarised across replicates
data_bd = db_bd %>%
  inner_join(db_sample, by = "sample_id") %>%
  group_by(sample_id, sample_name_bd, capture_id) %>%
  summarise(bd_qpcr_replicates = n(),
            bd_detected = any(detected, na.rm = TRUE),
            bd_mean_its1_copies_per_swab = mean(bd_its1_copies_per_swab, na.rm = TRUE),
            .groups = "drop")

# bring data queries together, select vars of interest, collect data
data_frog_sn = data_capture %>%
  left_join(data_bd, by = "capture_id") %>%
  left_join(data_env, by = "visit_id") %>%
  rename(survey_start_time = start_time,
         survey_end_time = end_time) %>%
  select(capture_id,
         taxon_capture,
         svl_mm,
         body_mass_g,
         sex,
         bd_qpcr_replicates,
         bd_detected,
         bd_mean_its1_copies_per_swab,
         body_temp_c,
         substrate_temp_c,
         microhabitat_type,
         microhabitat_wet,
         microhabitat_detailed,
         time_of_capture,
         timestamp_of_capture_utc,
         air_temp_c,
         air_time,
         water_temp_c,
         water_time,
         survey_start_time,
         start_timestamp_utc,
         survey_end_time,
         end_timestamp_utc,
         date,
         site_id,
         site,
         region,
         country,
         time_zone) %>%
  collect()

# export
write_csv(data_frog_sn, here("staging", paste0("frog_temp_sn_", today(), ".csv")))
```

# Example aligning frog temps with microclimate data

## collect water temp data
```{r}

# pull water temps (big data pulll)
water_temperature = db_ts_temp %>%
  left_join(db_sensor, by = "sensor_id") %>%
  left_join(db_logger, by = "logger_id") %>%
  inner_join(logger_site_date, by = "site_id") %>%
  filter(timestamp_utc >= ts_start_date,
         microhabitat == "water") %>%  # filter for water temp sensors only, multiple depths per site may be included
  pivot_wider(id_cols = c("site_id", "timestamp_utc"),
              names_from = c("microhabitat", "height_cm"),
              values_from = temperature_c_01_raw) %>%
  collect() %>%
  arrange(site_id,
          timestamp_utc)

# convert data_temperature to data.table
water_temp = as.data.table(water_temperature)

# include "sensor_time" column for nearest joins
water_temp_st = water_temp %>%
  mutate(sensor_time = timestamp_utc)
```

# rolling joins (data.table package) for aligning frog captures with water temperature
```{r}
# NEAREST in time
# align with site & nearest temp measurements in time
frog_temp_nearest = water_temp_st[data_frog, on = .(site_id, timestamp_utc = timestamp_of_capture_utc), roll = "nearest"] %>%
  rename(timestamp_of_cature_utc = timestamp_utc) %>%
  select(capture_id,
         any_of(colnames(db_capture)),
         everything())


# 15-day MEDIAN
# align with site & 15-day rolling median temp measurements in time
data_frog_dt = as.data.table(data_frog %>%
                               select(capture_id,
                                      site_id,
                                      timestamp_of_capture_utc))

frog_temp_15day_median = data_frog_dt[water_temp, on = .(site_id, timestamp_of_capture_utc = timestamp_utc), roll = days(-15), nomatch = NULL] %>%
  group_by(capture_id) %>%
  summarise(across(starts_with("water_"), ~ max(.x, na.rm = TRUE), .names = "{.col}_15d_med")) %>%
  right_join(data_frog, by = "capture_id") %>%
  select(capture_id,
         any_of(colnames(db_capture)),
         everything())
```