---
title: "frog_temp_microclimate_query"
output: html_document
---

# pull data for microclimate & frog body temp where we have both

Considerations
 - Loggers record at regular intevals generally not aligned with capture times. Fuzzy matching needed.
 - Not all captures have an associated time_of_capture, next best sit start_time and end_time of survey (usually ~1h range)
 - How do we want to 

```{r}

# librarian downloads, if not already downloaded, and reads in needed packages
librarian::shelf(tidyverse, dbplyr, here, RPostgres, DBI, RIBBiTR-BII/ribbitrrr)

# connect to database
dbcon = hopToDB("ribbitr")

```

# table pointers
```{r}
# survey data
db_bd = tbl(dbcon, Id("survey_data", "bd_qpcr_results"))
db_sample = tbl(dbcon, Id("survey_data", "sample"))
db_capture = tbl(dbcon, Id("survey_data", "capture"))
db_env = tbl(dbcon, Id("survey_data", "environmental"))
db_survey = tbl(dbcon, Id("survey_data", "survey"))
db_visit = tbl(dbcon, Id("survey_data", "visit"))
db_site = tbl(dbcon, Id("survey_data", "site"))
db_region = tbl(dbcon, Id("survey_data", "region"))
db_country = tbl(dbcon, Id("survey_data", "country"))

# microclimate data
db_ts_temp = tbl(dbcon, Id("microclimate_data", "ts_temperature"))
db_sensor = tbl(dbcon, Id("microclimate_data", "sensor"))
db_logger = tbl(dbcon, Id("microclimate_data", "logger"))
```

# joining data
```{r}
logger_site_date = db_ts_temp %>%
  left_join(db_sensor, by = "sensor_id") %>%
  left_join(db_logger, by = "logger_id") %>%
  group_by(site_id) %>%
  summarise(ts_start_date = date(min(timestamp_utc)),
            .groups = "drop")

# survey data
data_capture = db_capture %>%
  left_join(db_survey, by = "survey_id") %>%
  left_join(db_visit, by = "visit_id") %>%
  left_join(db_site, by = "site_id") %>%
  left_join(db_region, by = "region_id") %>%
  left_join(db_country, by = "country_id") %>%
  inner_join(logger_site_date, by = "site_id") %>%
  filter(date >= ts_start_date,
         !is.na(body_temp_c))

visit_ids = data_capture %>%
  select(visit_id) %>%
  distinct() %>%
  pull(visit_id)

# environmental surveys
data_env = db_env %>%
  left_join(db_survey, by = "survey_id") %>%
  left_join(db_visit, by = "visit_id") %>%
  filter(visit_id %in% visit_ids,
         !is.na(air_temp_c) | !is.na(water_temp_c)) %>%
  group_by(visit_id) %>%
  slice_min(order_by = tibble("air_time", "water_time"), n = 1) %>%
  ungroup() %>%
  select(visit_id,
         air_temp_c,
         air_time,
         water_temp_c,
         water_time)

# bd qpcr results
data_bd = db_bd %>%
  inner_join(db_sample, by = "sample_id") %>%
  group_by(sample_id, sample_name_bd, capture_id) %>%
  summarise(bd_qpcr_replicates = n(),
            bd_detected = any(detected, na.rm = TRUE),
            bd_mean_its1_copies_per_swab = mean(bd_its1_copies_per_swab, na.rm = TRUE),
            .groups = "drop")

data_frog = data_capture %>%
  left_join(data_env, by = "visit_id") %>%
  left_join(data_bd, by = "capture_id") %>%
  rename(survey_start_time = start_time,
         survey_end_time = end_time) %>%
  select(capture_id,
         taxon_capture,
         svl_mm,
         body_mass_g,
         sex,
         sample_name_bd,
         bd_qpcr_replicates,
         bd_detected,
         bd_mean_its1_copies_per_swab,
         body_temp_c,
         substrate_temp_c,
         microhabitat_type,
         microhabitat_wet,
         microhabitat_detailed,
         time_of_capture,
         air_temp_c,
         air_time,
         water_temp_c,
         water_time,
         survey_start_time,
         survey_end_time,
         date,
         site_id,
         site,
         region,
         country,
         time_zone) %>%
  collect()

```

# explore data
```{r}
frog_summary = data_frog %>%
  mutate(year = year(date)) %>%
  group_by(country, region, year) %>%
  count()

```

# demo of microclimate pairing script
```{r}

```