---
title: "penn_append"
format: html
editor: source
---
# Setup

## Load Packages

```{r}
librarian::shelf(tidyverse, dbplyr, here, janitor, lubridate, RPostgres, stringr, DBI, parsedate, uuid, hms, RIBBiTR-BII/ribbitrrr)
# librarian::shelf(RIBBiTR-BII/ribbitrrr, update_all = TRUE)

## Connect to DB
dbcon <- hopToDB()

## Pull metadata from database
mdc = tbl(dbcon, Id("public", "all_columns")) %>%
  filter(table_schema == "survey_data") %>%
  collect()

## Point to local data directory
ddir = Sys.getenv("data_dir")  # data directory
# Keeping .csv files in a central directory. Naming convention uses download date to distinguish batches.
wddir = here(ddir, "fulcrum", 'penn', '2024-11-18') # working data directory

```


# 0 - Data Plan

## Goals

- Integrate batches of Penn survey data with other RIBBiTR data (including other Penn data) for combined analysis
- Compare with existing data to avoid duplicating, and facilitate automated updating
- Quality control for red flag data (potential or known issues)
- Upload to Database in transaction
- log transaction

## Data chains (someday we can automate this...)

- capture / survey / visit / site / region / location
- aural / survey / visit / site / region / location
- ves / survey / visit / site / region / location

# 1 - Import data


## Pull dependent tables for each data chain

### use naming convention "db\_" to distinguish the source

```{r}
# pull relevant chain tables from DB
db_aural = tbl(dbcon, Id("survey_data", "aural"))
db_capture = tbl(dbcon, Id("survey_data", "capture"))
db_ves = tbl(dbcon, Id("survey_data", "ves"))
db_survey = tbl(dbcon, Id("survey_data", "survey"))
db_visit = tbl(dbcon, Id("survey_data", "visit"))
db_site = tbl(dbcon, Id("survey_data", "site"))
db_region = tbl(dbcon, Id("survey_data", "region"))
db_country = tbl(dbcon, Id("survey_data", "country"))

```


## Load most recent raw CSV Exports
- use naming convention "raw_" to distinguish source
- name related tables similarly for natural grouping

```{r}
# general survey table
raw_survey_info <- read_csv(here(wddir, "1_penn_surveyinformation", "1_penn_surveyinformation.csv"))

# aural tables
raw_aural_survey <- read_csv(here(wddir, "2_penn_acousticsurvey", "2_penn_acousticsurvey.csv"))
raw_aural_obs <- read_csv(here(wddir, "2_penn_acousticsurvey", "2_penn_acousticsurvey_acoustic_survey.csv"))

# ves tables
raw_ves_survey <- read_csv(here(wddir, "4_penn_visualencountersurvey", "4_penn_visualencountersurvey.csv"))
raw_ves_obs <- read_csv(here(wddir, "4_penn_visualencountersurvey", "4_penn_visualencountersurvey_visual_encounter_information.csv"))

# capture tables
raw_capture_survey <- read_csv(here(wddir, "5_penn_capturesurvey", "5_penn_capturesurvey.csv"))
raw_capture_obs <- read_csv(here(wddir, "5_penn_capturesurvey", "5_penn_capturesurvey_captured_amphibian_information.csv"))

raw_cmr <- read_csv(here(wddir, "supp_penn_cmrids", "supp_penn_cmrids.csv"))

# sample processing & obs tables
raw_sample_processing <- read_csv(here(wddir, "6_penn_sampleprocessing", "6_penn_sampleprocessing.csv"))
raw_sample_obs <- read_csv(here(wddir, "6_penn_sampleprocessing", "6_penn_sampleprocessing_amphibian_capture_survey_collections.csv"))

raw_edna_survey <- read_csv(here(wddir, "3_penn_ednasurvey", "3_penn_ednasurvey.csv"))
raw_edna_collection <- read_csv(here(wddir, "3_penn_ednasurvey", "3_penn_ednasurvey_edna_collection.csv"))
raw_edna_filtering <- read_csv(here(wddir, "3_penn_ednasurvey", "3_penn_ednasurvey_edna_collection_edna_filtering.csv"))

```
# 2 - Clean data

## check for null columns & duplicated rows
- purely informational, all action taken in next step

```{r}

null_columns = function(dataframe) {
  filtered = dataframe %>%
    select(where(~ all(is.na(.)))) %>%
    colnames()
}

# check null columns
(null_columns(raw_survey_info))
(null_columns(raw_aural_survey))
(null_columns(raw_aural_obs))
(null_columns(raw_ves_survey))
(null_columns(raw_ves_obs))
(null_columns(raw_capture_survey))
(null_columns(raw_capture_obs))
(null_columns(raw_cmr))
(null_columns(raw_sample_processing))
(null_columns(raw_sample_obs))

# check duplicated rows
any(duplicated(raw_survey_info))
any(duplicated(raw_aural_survey))
any(duplicated(raw_aural_obs))
any(duplicated(raw_ves_survey))
any(duplicated(raw_ves_obs))
any(duplicated(raw_capture_survey))
any(duplicated(raw_capture_obs))
any(duplicated(raw_cmr))
any(duplicated(raw_sample_processing))
any(duplicated(raw_sample_obs))

```

## Drop irrelevant & fix structural issues
- clean column names
- rename/mutate/drop specific columns
- Split, concatenate, coalesce related columns
- Convert to accurate data types/units/naming conventions
- drop unnecessary rows
- NA formatting

```{r}

# Fulcrum columns to drop. Best practice to drop by name
to_drop = c(
  "created_at",
  "updated_at",
  "created_by",
  "updated_by",
  "system_created_at",
  "system_updated_at",
  "version",
  "status",
  "project",
  "assigned_to",
  "latitude", # could this be useful?
  "longitude", # could this be useful?
  "geometry",
  "fulcrum_record_id",
  "site_other" # ignoring sites outside RIBBiTR... could include though.
)

calc_duration = function(start_time, end_time) {
  duration = if_else(end_time < start_time,
                         as_hms(86400) - start_time + end_time,
                         end_time - start_time)
  duration = duration / 60
  duration = str_remove(duration, " secs")
  duration = as.integer(round(as.numeric(duration), 0))
}

# Begin cleaning

clean_survey_info <- raw_survey_info %>%
  clean_names() %>% 
  rename(air_temp_c = air_temperature_c,
         air_time = air_temperature_measurement_time,
         water_temp_c = water_temperature_c,
         water_time = water_temperature_measurement_time,
         wind = wind_conditions,
         wind_speed_m_s = wind_speed_ms,
         sky = sky_conditions,
         relative_humidity_percent = humidity,
         pressure_psi = pressure,
         dissolved_o2_percent = dissolved_oxygen,
         tds_ppm = total_dissolved_solids,
         p_h = ph,
         comments_visit = sampling_event_comments,
         survey_info_id = fulcrum_id,
         time_of_day = survey_time) %>%
  unite(observers_survey, c("observers", "other_observers"), sep=",", remove = TRUE, na.rm = TRUE) %>%
  mutate(site = str_to_lower(site),
         site = str_replace_all(site, "-", "_"),
         site = str_replace_all(site, " ", "_"),
         time_of_day = str_to_lower(time_of_day),
         comments_visit = str_replace_all(comments_visit, "\\n", ". "),
         pressure_psi = ifelse(pressure_psi == 0, NA, pressure_psi),
         number_observers = as.integer(str_count(observers_survey, ",") + 1)) %>%
  select(-any_of(to_drop),
         -start_time,
         -end_time) %>%
  filter(!if_all(everything(), is.na))  # drop any rows all null

clean_aural_survey <- raw_aural_survey %>%
  clean_names() %>% 
  rename(comments_survey = acoustic_survey_comments,
         time_of_day = survey_time) %>%
  mutate(site = str_to_lower(site),
         site = str_replace_all(site, "-", "_"),
         site = str_replace_all(site, " ", "_"),
         time_of_day = str_to_lower(time_of_day),
         comments_survey = str_replace_all(comments_survey, "\\n", ". "),
         detection_type = "aural",
         created_at = with_tz(ymd_hms(created_at, tz = "UTC"), "America/New_York"),
         survey_date = as.Date(ifelse(hour(created_at) > 12, date(created_at), date(created_at) - days(1)), origin = "1970-01-01"),
         date = as.Date(ifelse(survey_date - date > 1, date, survey_date), origin = "1970-01-01")) %>%
  unite(observer_aural, c("observer", "observer_other"), sep=",", remove = TRUE, na.rm = TRUE) %>%
  select(-any_of(to_drop),
         -survey_date) %>%
  filter(!if_all(everything(), is.na),
         !is.na(site))  # drop any rows all null

clean_aural_obs <- raw_aural_obs %>%
  clean_names() %>%
  rename(species_aural = species_acoustic,
         species_aural_other = species_acoustic_other,
         comments_aural = acoustic_species_comments) %>%
  mutate(aural_id = fulcrum_id,
         species_aural = coalesce(species_aural, species_aural_other),
         species_aural = str_to_lower(str_replace_all(species_aural, " ", "_")),
         comments_aural = str_replace_all(comments_aural, "\\n", ". ")) %>%
  select(-any_of(to_drop),
         -species_aural_other) %>%
  filter(!if_all(everything(), is.na)) # drop any rows all null

clean_ves_survey <- raw_ves_survey %>%
  clean_names() %>% 
  rename(comments_survey = survey_description,
         time_of_day = survey_time) %>%
  mutate(site = str_to_lower(site),
         site = str_replace_all(site, "-", "_"),
         site = str_replace_all(site, " ", "_"),
         time_of_day = str_to_lower(time_of_day),
         comments_survey = str_replace_all(comments_survey, "\\n", ". "),
         detection_type = "visual",
         created_at = with_tz(ymd_hms(created_at, tz = "UTC"), "America/New_York"),
         survey_date = as.Date(ifelse(hour(created_at) > 12, date(created_at), date(created_at) - days(1)), origin = "1970-01-01"),
         date = as.Date(ifelse(survey_date - date > 1, date, survey_date), origin = "1970-01-01")) %>%
  unite(observer_ves, c("observer", "observer_other"), sep=",", remove = TRUE, na.rm = TRUE) %>%
  select(-any_of(to_drop),
         -survey_date) %>%
  filter(!if_all(everything(), is.na),
         !is.na(site))  # drop any rows all null

clean_ves_obs <- raw_ves_obs %>%
  clean_names() %>%
  mutate(ves_id = fulcrum_id,
         species_ves = coalesce(species_ves, species_ves_other),
         species_ves = str_to_lower(str_replace_all(species_ves, " ", "_")),
         comments_ves = str_replace_all(comments_ves, "\\n", ". ")) %>%
  select(-any_of(to_drop),
         -species_ves_other) %>%
  filter(!if_all(everything(), is.na))  # drop any rows all null

clean_capture_survey <- raw_capture_survey %>%
  clean_names() %>% 
  rename(comments_survey = survey_comments,
         time_of_day = survey_time) %>%
  mutate(site = str_to_lower(site),
         site = str_replace_all(site, "-", "_"),
         site = str_replace_all(site, " ", "_"),
         time_of_day = str_to_lower(time_of_day),
         comments_survey = str_replace_all(comments_survey, "\\n", ". "),
         detection_type = "capture",
         created_at = with_tz(ymd_hms(created_at, tz = "UTC"), "America/New_York"),
         survey_date = as.Date(ifelse(hour(created_at) > 12, date(created_at), date(created_at) - days(1)), origin = "1970-01-01"),
         date = as.Date(ifelse(survey_date - date > 1, date, survey_date), origin = "1970-01-01")) %>%
  unite(observer_capture, c("observer", "observer_other"), sep=",", remove = TRUE, na.rm = TRUE) %>%
  select(-any_of(to_drop),
         -survey_date) %>%
  filter(!if_all(everything(), is.na),
         !is.na(site))  # drop any rows all null

clean_capture_obs <- raw_capture_obs %>%
  clean_names() %>%
  rename(body_temp_c = body_temperature,
         comments_capture = amphibian_comments,
         capture_latitude = latitude,
         capture_longitude = longitude,
         photo_id = bag_photo) %>%
  mutate(capture_id = fulcrum_id,
         species_capture = coalesce(species_capture, species_capture_other),
         species_capture = str_to_lower(str_replace_all(species_capture, " ", "_")),
         comments_capture = str_replace_all(comments_capture, "\\n", ". "),
         bag_id = ifelse(bag_id == "BAG00", NA, bag_id),
         photo = !is.na(photo_id)) %>%
  select(-any_of(to_drop),
         -all_of(c("capture_status",
                   "species_capture_other",
                   "bag_photo_caption",
                   "bag_photo_url"))) %>% # capture status: was this individual successfully captured?
  filter(!if_all(everything(), is.na))  # drop any rows all null 

clean_cmr <- raw_cmr %>%
  clean_names() %>%
  rename(site = location) %>%
  mutate(site = str_to_lower(site),
         site = str_replace_all(site, "-", "_"),
         site = str_replace_all(site, " ", "_"),
         species = coalesce(species, species_other),
         cmr_id = coalesce(cmr_id, cmr_id_other)) %>%
  select(-any_of(to_drop),
         -species_other,
         -cmr_id_other) %>%
  filter(!if_all(everything(), is.na))  # drop any rows all null

clean_sample_processing <- raw_sample_processing %>%
  clean_names() %>%
  rename(site = location,
         comments_sample = survey_comments,
         time_of_day = survey_time) %>%
  mutate(site = str_to_lower(site),
         site = str_replace_all(site, "-", "_"),
         site = str_replace_all(site, " ", "_"),
         time_of_day = str_to_lower(time_of_day),
         created_at = with_tz(ymd_hms(created_at, tz = "UTC"), "America/New_York"),
         survey_date = as.Date(ifelse(hour(created_at) > 12, date(created_at), date(created_at) - days(1)), origin = "1970-01-01"),
         date = as.Date(ifelse(survey_date - date > 1, date, survey_date), origin = "1970-01-01")) %>%
  unite(processor, c("processor", "processor_other"), sep=",", remove = TRUE, na.rm = TRUE) %>%
  select(-any_of(to_drop),
         -survey_date,
         -location_other) %>%
  filter(!if_all(everything(), is.na)) # drop any rows all null

clean_sample_obs <- raw_sample_obs %>%
  clean_names() %>%
  rename(svl_mm = snout_vent_length,
         body_and_bag_mass_g = body_and_bag_mass,
         bag_mass_g = bag_mass,
         body_mass_g = body_mass,
         microbiome_swab_id = microbiome_id,
         comments_sample = capture_comments) %>%
  mutate(sex = str_to_lower(sex),
         sex_other = str_to_lower(sex_other),
         capture_type = str_to_lower(capture_type),
         life_stage = str_to_lower(life_stage),
         life_stage_other = str_to_lower(life_stage_other),
         species_capture = coalesce(species_capture, species_capture_other),
         species_capture = str_to_lower(str_replace_all(species_capture, " ", "_")),
         comments_sample = str_replace_all(comments_sample, "\\n", ". "),
         sample_id = fulcrum_id) %>%
  select(-any_of(to_drop),
         -fulcrum_record_id,
         -species_capture_other,
         -all_of(c("norep_bathvolume_50ml_acid_05ml",
                   "norep_bathvolume_100ml_acid_1ml",
                   "norep_bathvolume_200ml_acid_2ml",
                   "ul_of_norep_27mgml",
                   "ul_of_norep_135mgml",
                   "ul_of_norep_0675mgml",
                   "ul_of_norep_03375mgml",
                   "ul_of_norep_01688mgml",
                   "photo",
                   "photo_caption",
                   "photo_url",
                   "amp_ig_ne_injected",
                   "amp_ig_soak_volume",
                   "amp_ig_acidified",
                   "amp_ig_comments",
                   "bacterial_swab_id_foundations",
                   "bd_swab_id_foundations",
                   "microbiome_id_foundations",
                   "life_stage_other",
                   "sex_other"))) %>%
  filter(!if_all(everything(), is.na),
         !(species_capture == "negative_control"))

clean_edna_survey = raw_edna_survey %>%
  clean_names() %>%
  mutate(time_of_day = str_to_lower(survey_time))

clean_edna_collection = raw_edna_collection %>%
  clean_names()

clean_edna_filtering  = raw_edna_filtering %>%
  clean_names()


## Revisions: Handle special cases on a row-by-row basis
# (make repeatable & unique to observation, in case this exact observation(s) shows up again in a future data import)

# aquatic probe died, group returned later to measure. Comining two survey rows into one
row_a = clean_survey_info %>%
  filter(survey_info_id == "eea59aa0-4bae-4a1b-afa9-59c7285fbfe9")
row_b = clean_survey_info %>%
  filter(survey_info_id == "bbe6bef2-2ca6-43f3-b5e4-5db90f3ef82f")

revision_survey_info = row_a %>%
  mutate(observers_survey = paste0(row_a$observers_survey, ",", row_b$observers_survey),
         water_time = row_b$water_time,
         water_temp_c = row_b$water_temp_c,
         tds_ppm = row_b$tds_ppm,
         p_h = row_b$p_h)

clean_survey_info = clean_survey_info %>%
  filter(!survey_info_id %in% c("eea59aa0-4bae-4a1b-afa9-59c7285fbfe9",
                            "bbe6bef2-2ca6-43f3-b5e4-5db90f3ef82f")) %>%
  bind_rows(revision_survey_info)

clean_sample_obs$bag_id[clean_sample_obs$fulcrum_id == "9f72d0b0-9252-46e6-8391-8531174e1f97"] = "BAG68"
clean_sample_obs$bag_id[clean_sample_obs$fulcrum_id == "10e912f6-fdfb-4199-b7b0-aa137df728e4"] = "BAG70"
clean_sample_obs$bag_id[clean_sample_obs$fulcrum_id == "42c0d6a7-cd2b-442e-9fe7-d7e37363de14"] = "BAG32"
clean_sample_obs$bag_id[clean_sample_obs$fulcrum_id == "5a31383d-6210-42c7-8158-0f4818f0a66a"] = "BAG24"
clean_sample_obs$bag_id[clean_sample_obs$fulcrum_id == "c844de38-b38f-43d8-a2db-6d4373ffb275"] = "BAG34"
clean_sample_obs$bag_id[clean_sample_obs$fulcrum_id == "46c33c92-feb9-47ca-91d9-b519db773d69"] = "BAG64"
clean_sample_obs$bag_id[clean_sample_obs$fulcrum_id == "6d285d19-c288-456c-8dda-42ea35352dff"] = "BAG83"


clean_capture_obs$bag_id[clean_capture_obs$fulcrum_id == "78e8394d-60e4-4977-9f3b-9e57601b0980"] = "BAG10"
clean_capture_obs$bag_id[clean_capture_obs$fulcrum_id == "c513fba1-d5bb-41cb-add0-5330250c6d0f"] = "BAG06"
clean_capture_obs$bag_id[clean_capture_obs$fulcrum_id == "fb9736d2-d107-42a8-b5a0-09977b405010"] = "BAG18"
clean_capture_obs$bag_id[clean_capture_obs$fulcrum_id == "862a7efb-be07-40f8-ba1f-d0f4f90a8cd8"] = "BAG15"
clean_capture_obs$bag_id[clean_capture_obs$fulcrum_id == "4d5d73e8-847f-43e8-a8f1-db8c87371946"] = "BAG01"



```

# 3 Checks

## unify visit

```{r}
unique_visit <- bind_rows(clean_survey_info,
                          clean_aural_survey,
                          clean_capture_survey,
                          clean_edna_survey,
                          clean_ves_survey) %>% 
  select(date, site, time_of_day) %>%
  distinct()

unified_visit = clean_survey_info %>%
  full_join(unique_visit, by = c("site", "date", "time_of_day")) %>%
  filter(!is.na(site))

```

## gel visit
```{r}
gelled_visit = unified_visit %>%
  inner_join(db_site %>%
              select(tbl_keys("site", mdc)) %>%
              collect(),
            by = tbl_nkey("site", mdc)) %>%
  left_join(db_visit %>%
              select(tbl_keys("visit", mdc)) %>%
              collect(),
            by = tbl_nkey("visit", mdc)) %>%
  group_by_at(tbl_nkey("visit", mdc)) %>%
  mutate(visit_id = ifelse(is.na(visit_id), UUIDgenerate(), visit_id)) %>%
  ungroup()

```

## Generate stats for log
```{r}
dblog = list()
dblog$date_min = min(gelled_visit$date)
dblog$date_max = max(gelled_visit$date)
dblog$site = gelled_visit$site %>% unique()
dblog$site_id = gelled_visit$site_id %>% unique()
dblog$region = db_site %>%
  left_join(db_region, by = "region_id") %>%
  filter(site %in% dblog$sites) %>%
  pull(region) %>%
  unique()

```

## subset visit
```{r}
subset_visit = gelled_visit %>%
  select(any_of(colnames(db_visit))) %>%
  distinct() %>%
  filter(!is.na(visit_id))

compare_visit = compare_df_cols(db_visit %>%
                                  filter(FALSE) %>%
                                  collect(), subset_visit)



tray = compare_for_staging(db_visit %>%
                             filter(date >= dblog$date_min,
                                    site_id %in% dblog$site_id) %>%
                             collect(),
                           subset_visit, tbl_pkey('visit', mdc), return_all = TRUE, report = "visit")
final_visit = bind_rows(tray$insert,
                        tray$update)

# upsert visit?
x_visit = (nrow(final_visit) > 0)

```

## upsert visits

```{r}
if (x_visit) {
  temp_visit = stage_to_temp(dbcon, db_visit, final_visit)
  pointer = tbl(dbcon, temp_visit)
  db_visit = rows_upsert(db_visit, pointer, by="visit_id", in_place=TRUE)
}

```

## reload visits
```{r}
db_visit = tbl(dbcon, Id("survey_data", "visit"))
```

# One chain at a time

## Aural chain
- aural / survey / visit / site / region / location

### 3 - Unify aural data

```{r}
unified_aural = clean_aural_obs %>%
  full_join(clean_aural_survey, by = c("fulcrum_parent_id" = "fulcrum_id")) %>%
  select(-fulcrum_parent_id) %>%
  left_join(clean_survey_info, by = c("site", "date", "time_of_day")) %>%
  inner_join(db_site %>%
              select(tbl_keys("site", mdc)) %>%
              collect(),
            by = tbl_nkey("site", mdc)) %>%
  left_join(db_visit %>%
              select(tbl_keys("visit", mdc)) %>%
              collect(),
            by = tbl_nkey("visit", mdc)) %>%
  left_join(db_survey %>%
              select(tbl_keys("survey", mdc)) %>%
              filter(!is.na(visit_id)) %>%
              collect(),
            by = tbl_nkey("survey", mdc)) %>%
  filter(!is.na(site),
         !is.na(date))

```

### 4 - Gelled aural data

```{r}
gelled_aural = unified_aural %>%
  group_by_at(tbl_nkey("survey", mdc)) %>%
  mutate(start_time = as_hms(mean(start_time)),
         end_time = as_hms(mean(end_time)),
         duration_minutes = calc_duration(start_time, end_time),
         comments_survey = str_c(comments_survey, collapse = ", "),
         survey_id = ifelse(is.na(survey_id), first(fulcrum_id), survey_id),
         survey_id = ifelse(is.na(survey_id), UUIDgenerate(), survey_id)) %>%
  ungroup() %>%
  select(-fulcrum_id)

peace = gelled_aural %>%
  filter(survey_id == "21181bcc-4f52-44c0-a6e3-77fc83bd7a09")

```

### 5 - Subset aural tables

```{r}

# survey
subset_aural_survey = gelled_aural %>%
  select(any_of(colnames(db_survey))) %>%
  distinct() %>%
  filter(!is.na(survey_id))

# compare columns
compare_aural_survey = compare_df_cols(db_survey %>%
                                  filter(FALSE) %>%
                                  collect(), subset_aural_survey)

tray = compare_for_staging(db_survey %>% collect(), subset_aural_survey, tbl_pkey('survey', mdc), return_all = TRUE, report = "aural_survey")
final_aural_survey = bind_rows(tray$insert,
                               tray$update)

# aural
subset_aural = gelled_aural %>%
  select(any_of(colnames(db_aural))) %>%
  distinct() %>%
  filter(!is.na(aural_id))

# compare columns
compare_aural = compare_df_cols(db_aural %>%
                                  filter(FALSE) %>%
                                  collect(), subset_aural)

tray = compare_for_staging(db_aural %>% collect(), subset_aural, tbl_pkey('aural', mdc), return_all = TRUE, report = "aural")
final_aural = bind_rows(tray$insert,
                        tray$update)

# anything to upsert?
x_aural = (nrow(final_aural_survey) > 0) | (nrow(final_aural) > 0)

```

### Stage and commit aural tables

```{r}

if (x_aural) {
  # begin transaction temp
  dbBegin(dbcon)
  
  tryCatch(
    {
      temp_aural_survey = stage_to_temp(dbcon, db_survey, final_aural_survey)
      temp_aural = stage_to_temp(dbcon, db_aural, final_aural)
      
      
      pointer = tbl(dbcon, temp_aural_survey)
      rows_upsert(db_survey, pointer, by=tbl_nkey("survey", mdc), in_place=TRUE)
      
      pointer = tbl(dbcon, temp_aural)
      rows_upsert(db_aural, pointer, by=tbl_pkey("aural", mdc), in_place=TRUE)
      
      # Commit the transaction if successful
      dbCommit(dbcon)
      print("Transaction successful!")
      
    }, error = function(e) {
      # Rollback in case of error
      dbRollback(dbcon)
      message("Transaction failed: ", e$message)
    })
}
```


## VES chain
- ves / survey / visit / site / region / location

### Reload tables
```{r}
db_survey = tbl(dbcon, Id("survey_data", "survey"))
db_aural = tbl(dbcon, Id("survey_data", "aural"))
```



### Unify ves data
```{r}
unified_ves = clean_ves_obs %>%
  full_join(clean_ves_survey, by = c("fulcrum_parent_id" = "fulcrum_id")) %>%
  select(-fulcrum_parent_id) %>%
  left_join(clean_survey_info, by = c("site", "date", "time_of_day")) %>%
  inner_join(db_site %>%
              select(tbl_keys("site", mdc)) %>%
              collect(),
            by = tbl_nkey("site", mdc)) %>%
  left_join(db_visit %>%
              select(tbl_keys("visit", mdc)) %>%
              collect(),
            by = tbl_nkey("visit", mdc)) %>%
  left_join(db_survey %>%
              select(tbl_keys("survey", mdc)) %>%
              filter(!is.na(visit_id)) %>%
              collect(),
            by = tbl_nkey("survey", mdc)) %>%
  filter(!is.na(site),
         !is.na(date))

```

### Gel ves data
```{r}
gelled_ves = unified_ves %>%
  group_by_at(tbl_nkey("survey", mdc)) %>%
  mutate(start_time = as_hms(mean(start_time)),
         end_time = as_hms(mean(end_time)),
         duration_minutes = calc_duration(start_time, end_time),
         comments_survey = str_c(comments_survey, collapse = ", "),
         survey_id = ifelse(is.na(survey_id), first(fulcrum_id), survey_id),
         survey_id = ifelse(is.na(survey_id), UUIDgenerate(), survey_id)) %>%
  ungroup() %>%
  select(-fulcrum_id)
```

### 5 - Subset final ves tables

```{r}

# survey
subset_ves_survey = gelled_ves %>%
  select(any_of(colnames(db_survey))) %>%
  distinct() %>%
  filter(!is.na(survey_id))

# compare columns
compare_ves_survey = compare_df_cols(db_survey %>%
                                  filter(FALSE) %>%
                                  collect(), subset_ves_survey)

tray = compare_for_staging(db_survey %>% collect(), subset_ves_survey, tbl_pkey('survey', mdc), return_all = TRUE, report = "ves_survey")
final_ves_survey = bind_rows(tray$insert,
                             tray$update)


# aural
subset_ves = gelled_ves %>%
  select(any_of(colnames(db_ves))) %>%
  distinct() %>%
  filter(!is.na(ves_id))

# compare columns
compare_ves = compare_df_cols(db_ves %>%
                                  filter(FALSE) %>%
                                  collect(), subset_ves)

tray = compare_for_staging(db_ves %>% collect(), subset_ves, tbl_pkey('ves', mdc), return_all = TRUE, report = "ves")
final_ves = bind_rows(tray$insert,
                      tray$update)

x_ves = (nrow(final_ves_survey) > 0) | (nrow(final_ves) > 0)

```

### Stage and commit aural tables

```{r}
if (x_ves) {
  # begin transaction temp
  dbBegin(dbcon)
  
  tryCatch(
    {
      temp_ves_survey = stage_to_temp(dbcon, db_survey, final_ves_survey)
      temp_ves = stage_to_temp(dbcon, db_ves, final_ves)
      
      
      pointer = tbl(dbcon, temp_ves_survey)
      rows_upsert(db_survey, pointer, by=tbl_nkey("survey", mdc), in_place=TRUE)
      
      pointer = tbl(dbcon, temp_ves)
      rows_upsert(db_ves, pointer, by=tbl_pkey("ves", mdc), in_place=TRUE)
      
      # Commit the transaction if successful
      dbCommit(dbcon)
      print("Transaction successful!")
      
    }, error = function(e) {
      # Rollback in case of error
      dbRollback(dbcon)
      message("Transaction failed: ", e$message)
    })
}

```

## Reload tables
```{r}
db_survey = tbl(dbcon, Id("survey_data", "survey"))
db_ves = tbl(dbcon, Id("survey_data", "ves"))
```

## Capture chain
- capture / survey / visit / site / region / location

```{r}

unified_sample =  clean_sample_obs %>% 
  full_join(clean_sample_processing, by= c("fulcrum_parent_id" = "fulcrum_id")) %>%
  unite("comments_sample", c("comments_sample.x", "comments_sample.y"), sep=",", remove = TRUE, na.rm = TRUE) %>%
  select(-fulcrum_id,
         -fulcrum_parent_id)

unified_capture = clean_capture_obs %>%
  full_join(clean_capture_survey, by = c("fulcrum_parent_id" = "fulcrum_id")) %>%
  select(-fulcrum_parent_id) %>%
  mutate(temp_capture_id = fulcrum_id) %>%
  full_join(unified_sample, by=c("site", "date", "bag_id")) %>%
  mutate(species_capture = coalesce(species_capture.y, species_capture.x)) %>%
  select(-time_of_day.x,
         -time_of_day.y,
         -species_capture.x,
         -species_capture.y) %>%
  unite("comments_capture", c("comments_capture", "comments_sample"), sep=",", remove = TRUE, na.rm = TRUE) %>%
  left_join(clean_survey_info, by = c("site", "date")) %>%
  select(-fulcrum_id) %>%
  inner_join(db_site %>%
              select(tbl_keys("site", mdc)) %>%
              collect(),
            by = tbl_nkey("site", mdc)) %>%
  left_join(db_visit %>%
              select(tbl_keys("visit", mdc)) %>%
              collect(),
            by = tbl_nkey("visit", mdc)) %>%
  left_join(db_survey %>%
              select(tbl_keys("survey", mdc)) %>%
              collect(),
            by = tbl_nkey("survey", mdc)) %>%
  mutate(detection_type = "capture")

# 
# # QA/QC
# 
# # exploration
# sample_unmatched = unified_sample %>%
#   anti_join(unified_capture, by=c("site", "date", "bag_id")) %>%
#   # select(site, date, bag_id, species_capture, sample_id) %>%
#   mutate(capture_id = as.character(NA)) %>%
#   filter(!is.na(bag_id))
# 
# capture_unmatched = unified_capture %>%
#   anti_join(unified_sample, by=c("site", "date", "bag_id")) %>%
#   # select(site, date, bag_id, species_capture, capture_id) %>%
#   mutate(sample_id = as.character(NA)) %>%
#   filter(!is.na(bag_id))
# 
# sample_capture_unmatched = bind_rows(capture_unmatched, sample_unmatched) %>%
#   arrange(site, date, bag_id) %>%
#   group_by(site, date) %>%
#   mutate(temp_id = cur_group_id()) %>%
#   ungroup() %>%
#   select(site, date, bag_id, species_capture, capture_id, sample_id, start_time, end_time, observer_capture, processor, comments_capture, comments_sample,)
# 
# write_csv(sample_capture_unmatched, here("staging", "sample_campture_unmatched.csv"))
# 
# test_site = "rv_pond"
# test_bag = "BAG17"
# 
# peace = unified_sample %>%
#   filter(site == test_site,
#          bag_id == test_bag) %>%
#   arrange(bag_id)
# 
# train = unified_capture %>%
#   filter(site == test_site,
#          bag_id == test_bag) %>%
#   arrange(bag_id)

```

# Gelled capture

```{r}
gelled_capture = unified_capture %>%
  group_by_at(tbl_nkey("survey", mdc)) %>%
  mutate(start_time = as_hms(mean(start_time)),
         end_time = as_hms(mean(end_time)),
         duration_minutes = calc_duration(start_time, end_time),
         comments_survey = str_c(comments_survey, collapse = ", "),
         survey_id = ifelse(is.na(survey_id), first(temp_capture_id), survey_id),
         survey_id = ifelse(is.na(survey_id), UUIDgenerate(), survey_id)) %>%
  ungroup() %>%
  select(-temp_capture_id) %>%
  drop_na(visit_id)

```

### 5 - Subset final capture tables

```{r}
# survey
subset_capture_survey = gelled_capture %>%
  select(any_of(colnames(db_survey))) %>%
  distinct() %>%
  filter(!is.na(survey_id))

# compare columns
compare_capture_survey = compare_df_cols(db_survey %>%
                                  filter(FALSE) %>%
                                  collect(), subset_capture_survey)

tray = compare_for_staging(db_survey %>% collect(), subset_capture_survey, tbl_pkey('survey', mdc), return_all = TRUE, report = "capture_survey")

final_capture_survey = bind_rows(tray$insert,
                                 tray$update)

# capture
subset_capture = gelled_capture %>%
  select(any_of(colnames(db_capture))) %>%
  distinct() %>%
  filter(!is.na(capture_id))

# compare columns
compare_ves = compare_df_cols(db_ves %>%
                                  filter(FALSE) %>%
                                  collect(), subset_ves)


tray = compare_for_staging(db_capture %>% collect(), subset_capture, tbl_pkey('capture', mdc), return_all = TRUE, report = "capture")
final_capture = bind_rows(tray$insert,
                          tray$update)

x_capture = (nrow(final_capture_survey) > 0) | (nrow(final_capture) > 0)

```

### Stage and commit capture tables

```{r}
if (x_capture) {
  # begin transaction temp
  dbBegin(dbcon)
  
  tryCatch(
    {
      temp_capture_survey = stage_to_temp(dbcon, db_survey, final_capture_survey)
      temp_capture = stage_to_temp(dbcon, db_capture, final_capture)
      
      
      pointer = tbl(dbcon, temp_capture_survey)
      rows_upsert(db_survey, pointer, by=tbl_nkey("survey", mdc), in_place=TRUE)
      
      pointer = tbl(dbcon, temp_capture)
      rows_upsert(db_capture, pointer, by=tbl_pkey("capture", mdc), in_place=TRUE)
      
      # Commit the transaction if successful
      dbCommit(dbcon)
      print("Transaction successful!")
      
    }, error = function(e) {
      # Rollback in case of error
      dbRollback(dbcon)
      message("Transaction failed: ", e$message)
    })
}

```

## Reload tables
```{r}
db_survey = tbl(dbcon, Id("survey_data", "survey"))
db_capture = tbl(dbcon, Id("survey_data", "capture"))
```


## CMR table

```{r}

unified_cmr = clean_cmr %>%
  inner_join(db_site %>%
              select(tbl_keys("site", mdc)) %>%
              collect(),
            by = tbl_nkey("site", mdc)) %>%
  select(-site,
         -region_id)

```
